---
title: "House Price Prediction"
author: "Deepti Vijay Khandagale"
date: "5/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("G:/GMU/STAT-515 Applied Stattistics & Visulization of Analytics/project")
source("DataCleaning.R")
library(dplyr)
library(psych)
library(Hmisc)
library(ggplot2)
library(mice)
library(VIM)
library(missForest)
library(comprehenr)
library(ggpubr)
library(glmnet)
library(MASS)
library(randomForest)
library(caret)
library(Metrics)
```

## House Price Prediction

Ask a home buyer to describe their dream house, and they probably won't begin with the height of the basement ceiling or the proximity to an east-west railroad. But this playground competition's dataset proves that much more influences price negotiations than the number of bedrooms or a white-picket fence.

With 79 explanatory variables describing (almost) every aspect of residential homes in Ames, Iowa, this competition challenges you to predict the final price of each home.

```{r house price data , echo=TRUE}
train_house <- read.csv("train.csv")
head(train_house)
```


## Exploratory Data Analysis

### Missing Values and Imputation  

1. Dropping Columns with more than 40% missing values which are: 
  - PoolQC
  - MiscFeature
  - Alley
  - Fence
  - FireplaceQu

```{r pressure, echo=FALSE}
na_cols <- data.frame(colSums(is.na(train_house)))
na_cols <- cbind(Col_name = rownames(na_cols), na_cols)
colnames(na_cols ) <- c( 'Col_name' , 'Missing_Count')
na_cols <- na_cols[na_cols$Missing_Count != 0 , ]
na_cols <- na_cols[order(- na_cols$Missing_Count) , ]
rownames(na_cols) <- 1:nrow(na_cols)
na_cols$Total <- 1460
na_cols$perc_missing <- round(na_cols$Missing_Count / na_cols$Total , 3)
head(na_cols)
```

Other missing columns are 
```{r missing, echo=FALSE}
na_cols <- data.frame(colSums(is.na(train_house)))
na_cols <- cbind(Col_name = rownames(na_cols), na_cols)
colnames(na_cols ) <- c( 'Col_name' , 'Missing_Count')
na_cols <- na_cols[na_cols$Missing_Count != 0 , ]
na_cols <- na_cols[order(- na_cols$Missing_Count) , ]
rownames(na_cols) <- 1:nrow(na_cols)
na_cols$Total <- 1460
na_cols$perc_missing <- round(na_cols$Missing_Count / na_cols$Total , 3)
na_cols[na_cols$perc_missing < 0.4, ]
```
### Feature Engineering 

#### Categorical Variables 

There are couple of ways to handle the categorical variables. 
- Create Dummy Variables 
- Convert it into some ordinal data


#### Year of Construction


```{r YearBuild , echo=FALSE}

train_house$NumYear <-  (train_house$YearRemodAdd - train_house$YearBuilt)
train_house$YearBuilt <- as.factor(train_house$YearBuilt)
train_house$GarageYrBlt <- as.factor(train_house$GarageYrBlt)
train_house$YearRemodAdd <- as.factor(train_house$YearRemodAdd)
train_house$YrSold <- as.factor(train_house$YrSold)




yearBuilt <- ggplot(train_house , aes(x = YearBuilt , y = SalePrice)) + geom_boxplot() 
yearRemodled <- ggplot(train_house , aes(x = YearRemodAdd , y = SalePrice)) + geom_boxplot()
GarageYrBlt <- ggplot(train_house , aes(x = GarageYrBlt , y = SalePrice)) + geom_boxplot()
YrSold <- ggplot(train_house , aes(x = YrSold , y = SalePrice)) + geom_boxplot()


#scatter <- ggplot(train_house , aes(x = NumYear , y = SalePrice)) + geom_point()
ggarrange(yearBuilt, yearRemodled , GarageYrBlt , YrSold
          ,  ncol = 2, nrow = 2)

```


#### Basement

```{r Basement, echo= FALSE}


BsmtCond <- ggplot(subset(train_house , !is.na(train_house$BsmtCond)) 
       , aes(x = reorder(BsmtCond , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("BsmtCond")

BsmtExposure <- ggplot(subset(train_house , !is.na(train_house$BsmtExposure)) 
       , aes(x = reorder(BsmtExposure , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("BsmtExposure")

BsmtFinType1 <- ggplot(subset(train_house , !is.na(train_house$BsmtFinType1)) 
       , aes(x = reorder(BsmtFinType1 , SalePrice) , y = SalePrice)) + 
  geom_boxplot() + scale_x_discrete("BsmtFinType1")

BsmtFinType2 <- ggplot(subset(train_house , !is.na(train_house$BsmtFinType2)) 
       , aes(x = reorder(BsmtFinType2 , SalePrice) , y = SalePrice)) + 
  geom_boxplot() + scale_x_discrete("BsmtFinType2")

BsmtQual <- ggplot(subset(train_house , !is.na(train_house$BsmtQual)) 
       , aes(x = reorder(BsmtQual , SalePrice) , y = SalePrice)) + 
  geom_boxplot() + scale_x_discrete("BsmtQual")

ggarrange(BsmtCond, BsmtExposure , BsmtFinType1 , BsmtFinType2, BsmtQual
          ,  ncol = 3, nrow = 2)

```


#### Garage 

```{r Garage, echo= FALSE}


GarageCond <- ggplot(subset(train_house , !is.na(train_house$GarageCond)) 
       , aes(x = reorder(GarageCond , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("GarageCond")

GarageFinish <- ggplot(subset(train_house , !is.na(train_house$GarageFinish)) 
       , aes(x = reorder(GarageFinish , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("GarageFinish")

GarageQual <- ggplot(subset(train_house , !is.na(train_house$GarageQual)) 
       , aes(x = reorder(GarageQual , SalePrice) , y = SalePrice)) + 
  geom_boxplot() + scale_x_discrete("GarageQual")

GarageType <- ggplot(subset(train_house , !is.na(train_house$GarageType)) 
       , aes(x = reorder(GarageType , SalePrice) , y = SalePrice)) + 
  geom_boxplot() + scale_x_discrete("GarageType")


ggarrange( GarageCond , GarageFinish , GarageQual, GarageType
          ,  ncol = 2, nrow = 2)

# ggplot(train_house , aes(x = BsmtFinSF1 , y = SalePrice)) + geom_point()
# ggplot(train_house , aes(x = BsmtFinSF2 , y = SalePrice)) + geom_point()
```

#### Exterior


```{r Exterior, echo= FALSE}


ExterCond <- ggplot(subset(train_house , !is.na(train_house$ExterCond)) 
       , aes(x = reorder(ExterCond , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("ExterCond")

Exterior1st <- ggplot(subset(train_house , !is.na(train_house$Exterior1st)) 
       , aes(x = reorder(Exterior1st , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("Exterior1st") + theme(axis.text.x = element_text(angle = 90))

Exterior2nd <- ggplot(subset(train_house , !is.na(train_house$Exterior2nd)) 
       , aes(x = reorder(Exterior2nd , SalePrice) , y = SalePrice)) + 
  geom_boxplot() + scale_x_discrete("Exterior2nd") + theme(axis.text.x = element_text(angle = 90))

ExterQual <- ggplot(subset(train_house , !is.na(train_house$ExterQual)) 
       , aes(x = reorder(ExterQual , SalePrice) , y = SalePrice)) + 
  geom_boxplot() + scale_x_discrete("ExterQual")


ggarrange( ExterCond , Exterior1st , Exterior2nd, ExterQual
          ,  ncol = 2, nrow = 2)
```




#### Construction 

```{r Construction, echo= FALSE}


Foundation <- ggplot(subset(train_house , !is.na(train_house$Foundation)) 
       , aes(x = reorder(Foundation , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("Foundation")

LandContour <- ggplot(subset(train_house , !is.na(train_house$LandContour)) 
       , aes(x = reorder(LandContour , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("LandContour") 

LandSlope <- ggplot(subset(train_house , !is.na(train_house$LandSlope)) 
       , aes(x = reorder(LandSlope , SalePrice) , y = SalePrice)) + 
  geom_boxplot() + scale_x_discrete("LandSlope") 

BldgType <- ggplot(subset(train_house , !is.na(train_house$BldgType)) 
       , aes(x = reorder(BldgType , SalePrice) , y = SalePrice)) + 
  geom_boxplot() + scale_x_discrete("BldgType")

MasVnrType <- ggplot(subset(train_house , !is.na(train_house$MasVnrType)) 
       , aes(x = reorder(MasVnrType , SalePrice) , y = SalePrice)) + 
  geom_boxplot() + scale_x_discrete("MasVnrType")


LotShape <- ggplot(subset(train_house , !is.na(train_house$LotShape)) 
       , aes(x = reorder(LotShape , SalePrice) , y = SalePrice)) + 
  geom_boxplot() + scale_x_discrete("LotShape")

KitchenQual <- ggplot(subset(train_house , !is.na(train_house$KitchenQual)) 
       , aes(x = reorder(KitchenQual , SalePrice) , y = SalePrice)) + 
  geom_boxplot() + scale_x_discrete("KitchenQual")

RoofMatl <- ggplot(subset(train_house , !is.na(train_house$RoofMatl)) 
       , aes(x = reorder(RoofMatl , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("RoofMatl") + theme(axis.text.x = element_text(angle = 90))

RoofStyle <- ggplot(subset(train_house , !is.na(train_house$RoofStyle)) 
       , aes(x = reorder(RoofStyle , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("RoofStyle") + theme(axis.text.x = element_text(angle = 90))



ggarrange( Foundation , LandContour , LandSlope, BldgType , MasVnrType 
           ,LotShape , KitchenQual , RoofMatl , RoofStyle 
          ,  ncol = 3, nrow = 3)
```


#### Location 

```{r Location, echo= FALSE}


Condition1 <- ggplot(subset(train_house , !is.na(train_house$Condition1)) 
       , aes(x = reorder(Condition1 , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("Condition1") + theme(axis.text.x = element_text(angle = 90))

Condition2 <- ggplot(subset(train_house , !is.na(train_house$Condition2)) 
       , aes(x = reorder(Condition2 , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("Condition2") + theme(axis.text.x = element_text(angle = 90))

MSZoning <- ggplot(subset(train_house , !is.na(train_house$MSZoning)) 
       , aes(x = reorder(MSZoning , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("MSZoning") + theme(axis.text.x = element_text(angle = 90))

Neighborhood <- ggplot(subset(train_house , !is.na(train_house$Neighborhood)) 
       , aes(x = reorder(Neighborhood , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("Neighborhood") + theme(axis.text.x = element_text(angle = 90))


PavedDrive <- ggplot(subset(train_house , !is.na(train_house$PavedDrive)) 
       , aes(x = reorder(PavedDrive , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("PavedDrive") + theme(axis.text.x = element_text(angle = 90))



ggarrange(  Condition1 ,  Condition2 ,  MSZoning , Neighborhood  , PavedDrive
          ,  ncol = 3, nrow = 2)
```
#### Misc 

```{r Misc, echo= FALSE}


CentralAir <- ggplot(subset(train_house , !is.na(train_house$CentralAir)) 
       , aes(x = reorder(CentralAir , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("CentralAir") + theme(axis.text.x = element_text(angle = 90))

Electrical <- ggplot(subset(train_house , !is.na(train_house$Electrical)) 
       , aes(x = reorder(Electrical , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("Electrical") + theme(axis.text.x = element_text(angle = 90))

Functional <- ggplot(subset(train_house , !is.na(train_house$Functional)) 
       , aes(x = reorder(Functional , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("Functional") + theme(axis.text.x = element_text(angle = 90))

Heating <- ggplot(subset(train_house , !is.na(train_house$Heating)) 
       , aes(x = reorder(Heating , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("Heating") + theme(axis.text.x = element_text(angle = 90))


ggarrange(  CentralAir , Electrical , Functional , Heating
          ,  ncol = 2, nrow = 2)
```





#### Misc2 


```{r Misc2, echo= FALSE}


HeatingQC <- ggplot(subset(train_house , !is.na(train_house$HeatingQC)) 
       , aes(x = reorder(HeatingQC , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("HeatingQC") + theme(axis.text.x = element_text(angle = 90))

HouseStyle <- ggplot(subset(train_house , !is.na(train_house$HouseStyle)) 
       , aes(x = reorder(HouseStyle , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("HouseStyle") + theme(axis.text.x = element_text(angle = 90))

LotConfig <- ggplot(subset(train_house , !is.na(train_house$LotConfig)) 
       , aes(x = reorder(LotConfig , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("LotConfig") + theme(axis.text.x = element_text(angle = 90))

SaleCondition <- ggplot(subset(train_house , !is.na(train_house$SaleCondition)) 
       , aes(x = reorder(SaleCondition , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("SaleCondition") + theme(axis.text.x = element_text(angle = 90))


SaleType <- ggplot(subset(train_house , !is.na(train_house$SaleType)) 
       , aes(x = reorder(SaleType , SalePrice) , y = SalePrice)) +
  geom_boxplot() + scale_x_discrete("SaleType") + theme(axis.text.x = element_text(angle = 90))


ggarrange(  HeatingQC , HouseStyle , LotConfig , SaleCondition , SaleType
          ,  ncol = 3, nrow = 2)
```




### Checking the Linearity Assumptions.

- The Saleprice is rightly skewed which means that the assumption of linearity may not be true here. 
- Log of the Saleprice is normal distribution 


```{r SalePrice , echo=FALSE}

df_hist <- data.frame(train_house$SalePrice)
colnames(df_hist) <- c("SalePrice")
df_hist$SalePrice_log <- log(df_hist$SalePrice)

norm_hist <- ggplot(df_hist , aes(x = SalePrice)) + geom_histogram(bins = 50)
logHist <-   ggplot(df_hist , aes(x = SalePrice_log)) + geom_histogram(bins = 50)


ggarrange(  norm_hist , logHist,  ncol = 2, nrow = 1)

```

## Performing Data Cleaning 

```{r Data Cleaning, echo= TRUE}


train_house <- read.csv("train.csv")
train_house <- data_cleaning(train_house , thres = 0.4)
train_house <- feature_eng(train_house)

test_house <- read.csv("test.csv")
test_house <- TestDataPrep(test_house)

test_Saleprice <- read.csv("sample_submission.csv")
```

## Modeling

### 1. Multiple Linear Regression with Cross Validation

```{r Linear Model , echo= TRUE}
library(glmnet)
library(caret)

custom <- trainControl(method = "repeatedcv",
                       number = 10, 
                       repeats = 5, 
                       verboseIter = T)
lm <- train(log(SalePrice)~. , 
            train_house,
            method = 'lm',
            trControl = custom)

lm$results
lm
summary(lm)
plot(lm$finalModel)

```

### 2. Ridge Regression with Cross Validation

```{r Ridge , echo= TRUE}
library(glmnet)
library(caret)


set.seed(1234)
custom <- trainControl(method = "repeatedcv",
                       number = 10, 
                       repeats = 5, 
                       verboseIter = T)

ridge <- train(log(SalePrice)~. , 
            train_house,
            method = 'glmnet',
            tuneGrid = expand.grid(alpha = 0, 
                                   lambda = seq(0.0001,1,length = 5)),
            trControl = custom
  
)

plot(ridge)
ridge
plot(ridge$finalModel , xvar = 'lambda' , label = T)
plot(ridge$finalModel , xvar = 'dev' , label = T)
plot(varImp(ridge , scale = F) , top=30)
```

### 3. Lasso Regression with Cross Validation

```{r Lasso , echo= TRUE}
set.seed(1234)
custom <- trainControl(method = "repeatedcv",
                       number = 10, 
                       repeats = 5, 
                       verboseIter = T)

lasso <- train(log(SalePrice)~. , 
                    train_house,
                    method = 'glmnet',
                    tuneGrid = expand.grid(alpha = 1, 
                                           lambda = seq(0.0001,0.2,length = 5)),
                    trControl = custom
  
)

plot(lasso)
lasso
plot(lasso$finalModel , xvar = 'lambda' , label = T)
plot(lasso$finalModel , xvar = 'dev' , label = T)
plot(varImp(lasso , scale = FALSE) , top=30 )

```

# 4. Random Forest Method:
```{r}
set.seed(123)
RF = randomForest(log(train_house$SalePrice)~., data=train_house[-71],
                        mtry=20, ntree=500,importance=TRUE, na.action = na.omit)
RF

rmse(log(test_Saleprice$SalePrice), pred.RF )
mae(log(test_Saleprice$SalePrice), pred.RF)
mse(log(test_Saleprice$SalePrice), pred.RF)

```

#### Model Comparison

```{r modelComparison , echo= TRUE}

model_list <- list(LinearModel = lm , Ridge = ridge , Lasso = lasso)

res <- resamples(model_list)
summary(res)
bwplot(res)
```



#### Best Model

```{r Best Model , echo= TRUE}
lasso$bestTune

best <- lasso$finalModel
coef(best , s = lasso$bestTune$lambda)

```



# Prediction

# Multiple Linear Regression with cross validation

```{r}
pred.lm = predict(lm,newdata=test_house)

rmse(log(test_Saleprice$SalePrice), pred.lm )
mse(log(test_Saleprice$SalePrice), pred.lm )

```

#Ridge Regression with cross validation

```{r}
pred.ridge = predict(ridge,newdata=test_house)

rmse(log(test_Saleprice$SalePrice), pred.ridge)
mse(log(test_Saleprice$SalePrice), pred.ridge)
```


# Lasso Regression with cross Validation

```{r}
pred.lasso = predict(lasso,newdata=test_house)

rmse(log(test_Saleprice$SalePrice), pred.lasso )
mse(log(test_Saleprice$SalePrice), pred.lasso)
```


#Random Forest

```{r}
levels(test_house$Condition2) <- levels(train_house$Condition2)
levels(test_house$HouseStyle) <- levels(train_house$HouseStyle)
levels(test_house$Exterior1st) <- levels(train_house$Exterior1st)
levels(test_house$Exterior2nd) <- levels(train_house$Exterior2nd)
levels(test_house$RoofMatl) <- levels(train_house$RoofMatl)
levels(test_house$Heating) <- levels(train_house$Heating)
levels(test_house$Electrical) <- levels(train_house$Electrical)

pred.RF = predict(RF,newdata=test_house)

rmse(log(test_Saleprice$SalePrice), pred.RF )
mse(log(test_Saleprice$SalePrice), pred.RF)


```



<!-- #### Prediction -->

<!-- ```{r Predict , echo= TRUE} -->

<!-- p1 <- predict(best , train_house) -->
<!-- sqrt(mean((train_house$SalePrice - exp(p1))^2)) -->


<!-- ``` -->






<!-- #### RandomForest  -->

<!-- ```{r Rf , echo= TRUE} -->
<!-- library(randomForest) -->

<!-- rf <- randomForest(SalePrice ~ ., data = train_house, mtry = 3, ntree = 1000,  -->
<!--                          importance = TRUE, na.action = na.omit) -->

<!-- print(rf) -->
<!-- importance(rf) -->
<!-- ``` -->



#### RandomForest 

```{r stepwise, echo= TRUE}
model <- lm((log(SalePrice+1)) ~ . , data = train_house)
base_model <- lm((log(SalePrice+1)) ~ 1, data = train_house)
step = stepAIC(base_model ,direction = "forward", scope=list(lower=base_model,upper=formula(model)),trace = F)

par(mfrow=c(2,2))
plot(step)
step_error <- sqrt(mean(step$residuals^2))
```